DynaProg for Scala

### 2012-11-13 docs/roadmap/design.tex
NIts.

\newpage
\section{Architecture design and technical decisions}
% ------------------------------------------------------------------------------------------------
\subsection{Parsing grammar (ADP)}
{\color{red}
Borrowing from ADP XXX

>>>>

\newpage
\section{Architecture design and technical decisions}
% ------------------------------------------------------------------------------------------------
\subsection{User facing language requirements}
\quote{The Algebraic Dynamic Programming approach (ADP) introduces a conceptual splitting of a DP algorithm into a recognition and an evaluation phase. The evaluation phase is specified by an evaluation algebra, the recognition phase by a yield grammar. Each grammar can be combined with a variety of algebras to solve different but related problems, for which heretofore DP recurrences had to be developed independently. Grammar and algebra together describe a DP algorithm on a high level of abstraction, supporting the development of ideas and the comparison of algorithms. A notation for yield grammars is provided that serves as a domain-specific language.}

Given such an existing formalization \cite{adp} of dynamic programming on sequences, it seems natural to reuse it and possibly extend it to  other types of DP problems. In short, this framework allow the user to define a grammar using parsers, which are then run over an input string and produce intermediate results that are memoized into a table, when multiple solutions are possible, the user can define an aggregation function ($h$) to retain only desired candidates for further combination.

The benefits of ADP framework is that it does not constraint the result of the evaluation to be a single value, but could extend parsers to backtracking parsers or pretty-printers. Such functionality is easy to support in plain Scala, however, it could hampers the performance of the GPU implementation if we leave to the user the duty of representing and processing all the information within the parsers. Its major shortcoming in expressivity is that it only operates on strings, making only triangular-matrix problem easily representable.

The additional features we want to support are:\ol
\item Cyclic problems: are inherently very similar to string problems, except that the input is cyclic. To support such problem efficiently, we only need to mark the grammar cyclic such that it would apply on any unfolding of the cyclic input string.
\item Input pair algebra: the original ADP framework only support single 
\item Device restrictions and automatic backtracking: since efficient code has to be devised for C/CUDA implementation, we need to impose restrictions on the output that could be generated by the parsers. In particular, we would like to constraint the result to primary types (int, float, ...). However, on the other side, the backtracking information would be of primary interest for the DSL user, hence we need to automate the backtracking to both fulfill the goals of efficiency and usefulness of the device specific implementation.

\ole

\textbf{=> have a pair of subwords for multiple sequences}

XXX: explain ADP

Explain shortcomings

We want to offer the following features to the user:\ul
\item Support of cyclic problems (optionally with number of iterations)
\item Full features of ADP available in Scala, restriction on the CUDA available language
\item Windowing
\item Pair of inputs algebra (similar to SWat)
\ule

We define the aggregate function h as $List[T]\to List[T]$ and a special string emitter for C code that produces the initial element and a comparator between two (cost) elements.

% ------------------------------------------------------------------------------------------------
\subsection{Parsing grammar (ADP)}
{\color{red}
Borrowing from ADP XXX
